---
title: "Intro to Spatial Data in R"
format: html
---

## Downloading data
```{r}
knb_url <- "https://dev.nceas.ucsb.edu/knb/d1/mn/v2/object/urn%3Auuid%3Aaceaecb2-1ce0-4d41-a839-d3607d32bb58"

download.file(url = knb_url, destfile = '../shapefile_demo_data.zip')

unzip('../shapefile_demo_data.zip', exdir = '../data')

file.remove('../shapefile_demo_data.zip')
```

##Set up
```{r}
library(readr)
library(sf)
library(ggplot2)
library(leaflet)
library(scales)
library(ggspatial)
library(dplyr)
library(here)
```



#Read region data
```{r}
ak_regions <- read_sf(here("data", "ak_regions_simp.shp"))

plot(ak_regions)

class(ak_regions)
head(ak_regions)
glimpse(ak_regions)
st_crs(ak_regions)
```

##Transforming CRS

```{r}
ak_regions_3338 <- ak_regions %>% 
  st_transform(crs=3338)

plot(ak_regions_3338)

st_crs(ak_regions_3338)
```

##'sf' and the 'tidyverse'
```{r}
ak_regions_3338 %>% 
  select(region)
```

Filter for a specific region
```{r}
ak_regions_3338 %>% 
  filter(region == "Southeast")
```


## Spatial Joins
Read the population data
```{r}
pop <- read.csv(here("data", "alaska_population.csv"))
```

Convert pop data into a spatial object
```{r}
pop_4326 <- st_as_sf(pop,
                     coords= c("lng", "lat"),
                     crs = 4326,
                     remove= FALSE)
```

Joining population data with region data
```{r}
pop_joined <- st_join(pop_4326,
                      ak_regions_3338,
                      join = st_within)
```

Transform pop data into alaska alberts
```{r}
pop_3338 <- st_transform(pop_4326,
                         crs = 3338)
st_crs(pop_3338)
```

Join data with same crs
```{r}
pop_joined <- st_join(pop_3338,
                      ak_regions_3338,
                      join= st_within)
head(pop_joined)
```

Calculating total population by region

```{r}
pop_region <- pop_joined %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarise(total_pop= sum(population))

class(pop_region)
head(pop_region)
```


Joining Regions with total population by region

```{r}
pop_region_3338 <- left_join(ak_regions_3338,
                             pop_region,
                             by = "region")
plot(pop_region_3338["total_pop"])
```

Calculate total population by management area
```{r}
pop_mgmt_3338 <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarise(total_pop_mgmt = sum(total_pop),
            do_union = FALSE)

plot(pop_mgmt_3338["total_pop_mgmt"])
```

##Visualize with ggplot
```{r}
ggplot(pop_region_3338)+
  geom_sf(aes(fill = total_pop))+
  labs(fill= "Total Population")+
  scale_fill_continuous(low= "khaki",
                        high= "firebrick",
                        labels= comma)+
  theme_bw()
```








